---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity1 <- read.csv('activity.csv')
activity <- na.omit(activity1)
```

## What is mean total number of steps taken per day?
```{r}
library('knitr')
plotdata <- aggregate(activity[c("steps")], list(date = activity$date), sum)
kable(plotdata, format = "pandoc", digit = 2, align = 'l', caption = "Total Number of Steps Taken Per Day", col.names = c("Date", "Steps"))
```

```{r}
# code to plot histogram
hist(plotdata$steps, xlab = "Total number of steps taken each day", main = "Histogram of the Total Number of Steps Taken Each Day", col = "Blue")
```

```{r}
#code to calculate the mean and median of the total number of steps taken each day
means_steps <- format(mean(plotdata$steps), digits = 1)
median_steps <- median(plotdata$steps)
```
#### The mean of the total number of steps taken each day is `r means_steps`
#### The median of the total number of steps taken each day is `r median_steps`

## What is the average daily activity pattern?
```{r}
timeseries <- aggregate(activity[c("steps")], list(interval = activity$interval), mean)
plot(timeseries$interval, timeseries$steps, type="l", xlab = "Interval", ylab = "Average Number of Steps Taken", main = "Average Daily Activity Pattern")

# find which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
max_steps_interval = timeseries$interval[which.max(timeseries$steps)]
```

#### Interval, on average across all the days in the dataset that contains the maximum number of steps is `r max_steps_interval`

## Imputing missing values
```{r}
missing_values <- sum(is.na(activity1))
```
#### Number of missing values = `r missing_values`
```{r}
for (i in 1:nrow(activity1)) {
    if (is.na (activity1$steps[i])) {
         activity1$steps[i] <- subset(timeseries, timeseries$interval == activity1$interval[i])$steps
    }
}
plotdata <- aggregate(activity1[c("steps")], list(date = activity1$date), sum)
hist(plotdata$steps, xlab = "Total number of steps taken each day", main = "Histogram of the Total Number of Steps Taken Each Day", col = "Blue")
```

```{r}
#code to calculate the mean and median of the total number of steps taken each day
means_steps <- format(mean(plotdata$steps), digits = 1)
median_steps <- format(median(plotdata$steps), digits = 1)
```
#### The mean of the total number of steps taken each day is `r means_steps`
#### The median of the total number of steps taken each day is `r median_steps`
#### Only the median is different from the first part of this assignment. Imputing missing values has minimal impact on the results. 

## Are there differences in activity patterns between weekdays and weekends?
